# Desrired workplace and residential census tracts for Brooklyn, NY
## Folks that live and work within brooklyn

## Pre-process data
Import required libraries
```{r, message=FALSE}
library(sf)
library(geojsonsf)
library(tidyverse)
library(tmap)
```

Define places of interest
```{r, message=FALSE}
boro_name = "Brooklyn"
county_code = "047"
```

Import required data
```{r, message=FALSE}
options(scipen = 999)
nyc_tract_borders <- geojson_sf('./data/nyc_2010_tract_borders.geojson')
ny_origin_dest <- read_csv("./data/ny_od_main_JT00_2019.csv")
```

Reduce NYC tract borders to Brooklyn
```{r, message = FALSE}
bk_tract_borders <- nyc_tract_borders %>%
  filter(BoroName == boro_name) %>%
  select(c("CT2010", "geometry"))
```

Reduce NYS origin destination data to only census tracts within Brooklyn
```{r, message = FALSE}
bk_ods <- ny_origin_dest %>%
  filter(
    str_sub(as.character(w_geocode), 3, 5) == county_code &
      str_sub(as.character(h_geocode), 3, 5) == county_code
    ) %>%
  mutate(w_tract = str_sub(as.character(w_geocode), 6, 11)) %>%
  mutate(h_tract = str_sub(as.character(h_geocode), 6, 11)) %>%
  select(h_tract, w_tract, S000) %>%
  mutate(od = str_c(h_tract, w_tract)) %>%
  group_by(od) %>%
  summarise(
    h_tract,
    w_tract,
    S000 = sum(S000),
    ) %>%
  unique()
```

Define most popular work tracts
```{r, message=FALSE}
pop_work_tracts <- bk_ods %>%
  group_by(w_tract) %>%
  summarise(
    w_tract,
    S000 = sum(S000)
  ) %>%
  unique() %>%
  left_join(bk_tract_borders, c("w_tract"= "CT2010")) %>%
  st_as_sf()
```

Define most popular home tracts
```{r, message=FALSE}
pop_home_tracts <- bk_ods %>%
  group_by(h_tract) %>%
  summarise(
    h_tract,
    S000 = sum(S000)
  ) %>%
  unique() %>%
  left_join(bk_tract_borders, c("h_tract"= "CT2010")) %>%
  st_as_sf()
```
### Utility functions

## Most popular tracts for folks to work and live within Brooklyn
```{r,  out.width="100%", fig.align='center'}
pop_work_map <- tm_shape(pop_work_tracts) +
  tm_polygons(
    col = "S000",
    style = "jenks",
    title = "Employers",
  ) + 
  tm_layout(
    title = "Work Tracts",
    legend.outside = TRUE
  )
pop_home_map <- tm_shape(pop_home_tracts) +
  tm_polygons(
    col = "S000",
    style = "jenks",
    title = "Homes",
  ) + 
  tm_layout(
    title = "Residential Tracts",
    legend.outside = TRUE
  )

tmap_arrange(pop_work_map, pop_home_map)
```

## Most popular tracts for folks to work and live within the same tract

## Most popular desire lines for folks to commute from one to tract to another