# Origins and Destinations Brooklyn NTAs in 2002 and 2010, uncorrected for tract changes

## Preprocess Data

Import necessary files
```{r, message=FALSE }
library(sf)
library(geojsonsf)
library(tidyverse)
library(tmap)
```

Define places of interest
```{r, message=FALSE}
boro_name = "Brooklyn"
county_code = "047"
parks <- "BK99"
```

Import required data
```{r, message=FALSE}
options(scipen = 999)
nyc_nta_borders <- geojson_sf('./data/nyc_2010_nta_borders.geojson')
nyc_tract_nta_equiv <- readxl::read_xlsx('./data/nyc_2010_census_tract_nta_equiv.xlsx')
ny_od_02 <- read_csv("./data/ny_od_main_JT00_2002.csv")
ny_od_19 <- read_csv("./data/ny_od_main_JT00_2019.csv")
```

Create equivalency for Brooklyn tracts and ntas
```{r, message = FALSE}
bk_tract_nta_equiv <- nyc_tract_nta_equiv %>%
  filter(borough_name == boro_name & nta_code != parks) %>%
  mutate(tract = `census_tract`) %>%
  select("tract", "nta_code")
```

Reduce NYC nta borders to Brooklyn
```{r, message = FALSE}
bk_nta_borders <- nyc_nta_borders %>%
  filter(BoroName == boro_name & NTACode != parks) %>%
  select(c("NTACode", "geometry"))
```

Find points on surface of Brooklyn ntas
```{r, message=FALSE}
bk_nta_points <- bk_nta_borders %>%
  mutate(geometry = st_point_on_surface(geometry))
```
Reduce 2019 NYS origin destination data to trips that start and end in Brooklyn
```{r, message = FALSE}
bk_nta_od_19 <- ny_od_19 %>%
  # trip starts and ends in Brooklyn
  filter(
    str_sub(as.character(w_geocode), 3, 5) == county_code &
      str_sub(as.character(h_geocode), 3, 5) == county_code
    ) %>%
  # extract the tract. it's already known to be in Brooklyn
  mutate(w_tract = str_sub(as.character(w_geocode), 6, 11)) %>%
  mutate(h_tract = str_sub(as.character(h_geocode), 6, 11)) %>%
  # only interested in home and work tracts, and the number of total trips
  select(h_tract, w_tract, S000) %>%
  # map home tract to NTA
  left_join(bk_tract_nta_equiv, c("h_tract" = "tract")) %>%
  rename(h_nta_code = nta_code) %>%
  # map work tract to NTA
  left_join(bk_tract_nta_equiv, c("w_tract" = "tract")) %>%
  rename(w_nta_code = nta_code) %>%
  # remove parks
  filter(h_nta_code != parks & w_nta_code != parks) %>%
  # create key to identify trip route
  mutate(od = str_c(h_nta_code, w_nta_code)) %>%
  # count number of trips for each route, removing duplicate entries
  group_by(od) %>%
  summarise(
    h_nta_code,
    w_nta_code,
    S000 = sum(S000),
    ) %>%
  unique()
```

Reduce 2002 NYS origin destination data to trips that start and end in Brooklyn
```{r, message = FALSE}
bk_nta_od_02 <- ny_od_02 %>%
  # trip starts and ends in Brooklyn
  filter(
    str_sub(as.character(w_geocode), 3, 5) == county_code &
      str_sub(as.character(h_geocode), 3, 5) == county_code
    ) %>%
  # extract the tract. it's already known to be in Brooklyn
  mutate(w_tract = str_sub(as.character(w_geocode), 6, 11)) %>%
  mutate(h_tract = str_sub(as.character(h_geocode), 6, 11)) %>%
  # only interested in home and work tracts, and the number of total trips
  select(h_tract, w_tract, S000) %>%
  # map home tract to NTA
  left_join(bk_tract_nta_equiv, c("h_tract" = "tract")) %>%
  rename(h_nta_code = nta_code) %>%
  # map work tract to NTA
  left_join(bk_tract_nta_equiv, c("w_tract" = "tract")) %>%
  rename(w_nta_code = nta_code) %>%
  # remove parks
  filter(h_nta_code != parks & w_nta_code != parks) %>%
  # create key to identify trip route
  mutate(od = str_c(h_nta_code, w_nta_code)) %>%
  # count number of trips for each route, removing duplicate entries
  group_by(od) %>%
  summarise(
    h_nta_code,
    w_nta_code,
    S000 = sum(S000),
    ) %>%
  unique()
```

Define most popular work ntas in 2019
```{r, message=FALSE}
bk_nta_dest_19 <- bk_nta_od_19 %>%
  group_by(w_nta_code) %>%
  summarise(
    w_nta_code,
    S000 = sum(S000)
  ) %>%
  unique() %>%
  left_join(bk_nta_borders, c("w_nta_code"= "NTACode")) %>%
  st_as_sf()
```
Define most popular work ntas in 2002
```{r, message=FALSE}
bk_nta_dest_02 <- bk_nta_od_02 %>%
  group_by(w_nta_code) %>%
  summarise(
    w_nta_code,
    S000 = sum(S000)
  ) %>%
  unique() %>%
  left_join(bk_nta_borders, c("w_nta_code"= "NTACode")) %>%
  st_as_sf()
```

Define most popular home ntas in 2019
```{r, message=FALSE}
bk_nta_origin_19 <- bk_nta_od_19 %>%
  group_by(h_nta_code) %>%
  summarise(
    h_nta_code,
    S000 = sum(S000)
  ) %>%
  unique() %>%
  left_join(bk_nta_borders, c("h_nta_code"= "NTACode")) %>%
  st_as_sf()
```

Define most popular home ntas in 2002
```{r, message=FALSE}
bk_nta_origin_02 <- bk_nta_od_02 %>%
  group_by(h_nta_code) %>%
  summarise(
    h_nta_code,
    S000 = sum(S000)
  ) %>%
  unique() %>%
  left_join(bk_nta_borders, c("h_nta_code"= "NTACode")) %>%
  st_as_sf()
```
## Most popular tracts to work in Brooklyn across 2002 and 2019
```{r,  out.width="100%", fig.align='center'}
bk_dest_dist_02 <- tm_shape(bk_nta_dest_02) +
  tm_polygons(
    col = "S000",
    style = "jenks",
    title = "2002",
  ) + 
  tm_layout(
    title = "Work Tracts",
    legend.outside = TRUE
  )
bk_dest_dist_19 <- tm_shape(bk_nta_dest_19) +
  tm_polygons(
    col = "S000",
    style = "jenks",
    title = "2019",
  ) + 
  tm_layout(
    title = "Work Tracts",
    legend.outside = TRUE
  )

tmap_arrange(bk_dest_dist_02, bk_dest_dist_19)
```

## Most popular tracts to work in Brooklyn across 2002 and 2019
```{r,  out.width="100%", fig.align='center'}
bk_origin_dist_02 <- tm_shape(bk_nta_origin_02) +
  tm_polygons(
    col = "S000",
    style = "jenks",
    title = "2002",
  ) + 
  tm_layout(
    title = "Home Tracts",
    legend.outside = TRUE
  )
bk_origin_dist_19 <- tm_shape(bk_nta_origin_19) +
  tm_polygons(
    col = "S000",
    style = "jenks",
    title = "2019",
  ) + 
  tm_layout(
    title = "Home Tracts",
    legend.outside = TRUE
  )

tmap_arrange(bk_origin_dist_02, bk_origin_dist_19)
```