---
output:
  pdf_document: default
  html_document: default
---
## Necessary libraries
```{r message = FALSE}
library(tidyverse)
library(geojsonsf)
library(sf)
library(tmap)
options(scipen = 999)
```
# Outline
- Changes to the topic
  - Shift from trying to predict changes in dynamics to instead focusing on how much of an impact each travel mode makes
- Changes to the timeline
  - Need to create a timeline
  - retroactively mark when progress has been made. Look forward to what progress still needs to be done.
- Review of the original timeline: 
  - more detailed information on what has changed for the project
- A more detailed version of the methods to be applied.

## Questions
- corrections for changes in census tracts
### Updated approach
- Network auto-correlation of OD desire lines for 2019
- Network auto-correlation of MTA system
- Network auto-correlation of Highway system
- Network correlation of contiguous NTAs
- Regression to compare the MTA and Highway systems, to see which has the greatest effect
- Is there are way to measure the shift in the networks?

Started with interest in Triboro line ridership
- Interest in the four boroughs that are severed by the subway system
- Manhattan overwhelms the travel patterns throughout the boroughs of interest
- Insufficient data to track the total ridership and the means 
  - Travel surveys did not offer the level of detail or confidence
- Shifted to looking specifically at work commuting patterns within a single borough
- Focused on what correlation each transporation system has with the OD patterns

## Data sources
- Origin and Destination data for NYS in 2019 at the block level
- NYC Census tract borders
- Borders of NYC boroughs
- Equivalency of Neighborhood Tabulation Areas
- Subway routes
  - Good for data exploration
  - Will need to be suplemented with a routing service
- Arterials and Major Roads
  - Good for data exploration
  - Will need to be supplemented with a routing service

## Visualize the census tracts and ntas, providing context for their size
```{r, message = FALSE}
bois_names <- c("Manhattan", "Bronx", "Brooklyn", "Queens")
bois_census_tract_borders <- geojson_sf('./data/nyc_2010_census_tract_borders.geojson') %>%
  dplyr::filter(BoroName %in% bois_names)
  
tmap::tm_shape(bois_census_tract_borders) + 
  tmap::tm_polygons(
    col = "BoroName",
    title = "Borough"
  )
```

### Estimate the average distance from the center to the edge of tract
```{r, message = FALSE}
avg_tract_area <- sum(bois_census_tract_borders$Shape__Area)/ length(bois_census_tract_borders$Shape__Area)
avg_tract_radius <- sqrt(avg_tract_area/pi)
print(avg_tract_radius)
```

## Visualize the ntas
```{r, message = FALSE}
bois_nta_borders <- geojson_sf('./data/nyc_2010_nta_borders.geojson') %>%
  dplyr::filter(BoroName %in% bois_names)

tmap::tm_shape(bois_nta_borders) +
  tmap::tm_polygons(
    col = "BoroName",
    title = "Borough"
  )
```
```{r, message = FALSE}
avg_nta_area <- sum(bois_nta_borders$Shape__Area) / length(bois_nta_borders$Shape__Area)
avg_nta_radius <- sqrt(avg_nta_area/pi)
print(avg_nta_radius)
```
## Demonstrate the disparity between Manhattan and the other boroughs of interest 

### Create map between census tracts and ntas
Specify county because census tracts may not be unique across counties
```{r, message = FALSE}
bois_county_tract_nta_equiv <- readxl::read_xlsx('./data/nyc_2010_census_tract_nta_equiv.xlsx') %>%
  filter(borough_name %in% bois_names) %>%
  mutate(county_tract = str_c(`county_code`, `census_tract`)) %>%
  select("county_tract", "nta_code")
```

Reduce NYS origin destination data to only ntas of interest
```{r, message = FALSE}
bois_county_codes = c("061", "005", "047", "081") # Manhattan, Bronx, Brooklyn, Queens
bois_nta_ods <- read_csv('./data/ny_od_main_JT00_2019.csv') %>%
  # Select only tracts within the boroughs of interest
  dplyr::filter(
    stringr::str_sub(as.character(w_geocode), 3, 5) %in% bois_county_codes &
      stringr::str_sub(as.character(h_geocode), 3, 5) %in% bois_county_codes
    ) %>%
  # Create fields specifically for home counties and tracts
  mutate(w_county_tract = str_sub(as.character(w_geocode), 3, 11)) %>%
  mutate(h_county_tract = str_sub(as.character(h_geocode), 3, 11)) %>%
  # Narrow table down to tracts and all jobs
  select(h_county_tract, w_county_tract, S000) %>%
  # Relate tracts with ntas
  left_join(bois_county_tract_nta_equiv, c("h_county_tract" = "county_tract")) %>%
  rename(h_nta_code = nta_code) %>%
  left_join(bois_county_tract_nta_equiv, c("w_county_tract" = "county_tract")) %>%
  rename(w_nta_code = nta_code) %>%
  # Remove trips within the same nta, only inter-nta trips are of interest
  filter(w_nta_code != h_nta_code) %>%
  # Label trips based on home and work ntas
  mutate(od = str_c(h_nta_code, w_nta_code)) %>%
  # Count the number of trips made between these ntas and in this direction 
  group_by(od) %>%
  summarise(
    h_nta_code,
    w_nta_code,
    S000 = sum(S000),
    ) %>%
  # Remove duplicate entries
  unique()
```

