# Desire Lines between NYC NTAs

Import required libraries
```{r, message=FALSE }
library(sf)
library(geojsonsf)
library(tidyverse)
library(tmap)
```

Import data from files
- NYC NTA polygons
- NYC NTA to tract equivalence
- NYS OD 
```{r, message=FALSE}
options(scipen = 999)
nyc_nta_geo <- geojson_sf('./data/nyc_nta_2010.geojson')
nyc_tract_nta_equiv <- readxl::read_xlsx('./data/nyc_2010_census_tract_nta_equiv.xlsx')
ny_origin_dest <- read_csv("./data/ny_od_main_JT00_2019.csv")
```

Define places of interest
```{r, message = FALSE}
boros <- c("Bronx", "Brooklyn", "Manhattan", "Queens")
county_codes <- c("005", "047", "061", "081")
parks <- c("BX10", "BX99", "BK99", "MN99", "QN99")
```

Create equivalency for counties and tracts with NTAs
```{r, message = FALSE}
county_tract_nta_equiv <- readxl::read_xlsx('./data/nyc_2010_census_tract_nta_equiv.xlsx') %>%
  filter(!(nta_code %in% parks)) %>%
  mutate(county_tract = str_c(`county_code`, `census_tract`)) %>%
  select("county_tract", "nta_code")
```

Reduce NYS origin destination data to only ntas of interest
```{r, message = FALSE}
nta_ods <- ny_origin_dest %>%
  filter(
    str_sub(as.character(w_geocode), 3, 5) %in% county_codes &
      str_sub(as.character(h_geocode), 3, 5) %in% county_codes
    ) %>%
  mutate(w_county_tract = str_sub(as.character(w_geocode), 3, 11)) %>%
  mutate(h_county_tract = str_sub(as.character(h_geocode), 3, 11)) %>%
  select(h_county_tract, w_county_tract, S000) %>%
  left_join(county_tract_nta_equiv, c("h_county_tract" = "county_tract")) %>%
  rename(h_nta_code = nta_code) %>%
  left_join(county_tract_nta_equiv, c("w_county_tract" = "county_tract")) %>%
  rename(w_nta_code = nta_code) %>%
  filter(w_nta_code != h_nta_code) %>%
  mutate(od = str_c(h_nta_code, w_nta_code)) %>%
  group_by(od) %>%
  summarise(
    h_nta_code,
    w_nta_code,
    S000 = sum(S000),
    ) %>%
  unique()
```

Reduce NYC geography to NTAs of interest
```{r, message= FALSE}
nta_polys <- nyc_nta_geo %>%
  filter(CountyFIPS %in% county_codes) %>%
  filter(!(NTACode %in% parks)) %>%
  select(BoroName, NTACode, geometry)
```

Find points in NTA polygons
```{r, message = FALSE}
nta_points <- nta_polys %>%
  mutate(geometry = st_point_on_surface(geometry))
```

Define desire lines for origins and destinations
```{r, message=FALSE}
nta_od_lines <- nta_ods %>%
  left_join(nta_points, c("h_nta_code" = "NTACode")) %>%
  rename(h_geometry = geometry) %>%
  left_join(nta_points, c("w_nta_code" = "NTACode")) %>%
  rename(w_geometry = geometry) %>%
  mutate(geometry = st_union(h_geometry, w_geometry)) %>%
  mutate(geometry = st_cast(geometry, "LINESTRING")) %>%
  select("od", "S000", "geometry") %>%
  st_as_sf()
```

Map of desire lines between NTAs
```{r, out.width="100%", fig.align='center'}
tm_shape(nta_polys) +
  tm_polygons(
    col = "BoroName",
  ) +
  tm_shape(nta_od_lines) +
    tm_lines(
      col = "#212121",
      lwd = "S000",
      )
```